---
output: github_document
---

<!-- badges: start -->

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/aweek)](https://cran.r-project.org/package=aweek) 
[![Codecov test coverage](https://codecov.io/gh/zkamvar/aweek/branch/master/graph/badge.svg)](https://codecov.io/gh/zkamvar/aweek?branch=master) 
[![Travis build status](https://travis-ci.org/zkamvar/aweek.svg?branch=master)](https://travis-ci.org/zkamvar/aweek)

<!-- badges: end -->

# Welcome to the *aweek* package!

This package will convert dates to
[epiweeks](http://www.cmmcp.org/epiweek.htm),
[isoweeks](http://en.wikipedia.org/wiki/ISO_week_date), and all others in
between with minimal overhead. Other packages such as
[lubridate](https://github.com/tidyverse/lubridate),
[ISOweek](https://cran.r-project.org/package=ISOweek), and
[surveillance](http://surveillance.r-forge.r-project.org/) have the ability to
define ISOweeks and/or epiweeks have the ability to define ISOweeks and/or
epiweeks, the ability to easily switch between day and week intervals is only
available for ISOweeks. However, the ability to easily switch between day and
week intervals is only available for ISOweeks and all of the packages above
have dependencies that require compiled code.


```{r epiweek}

library(aweek)
set.seed(2019-02-26)
onset <- Sys.Date() + sort(sample(-6:7, 20, replace = TRUE))
onset
date2week(onset, week_start = 1) # ISO weeks starting on Monday (default)
date2week(onset, week_start = 7) # EPI week starting on Sunday
date2week(onset, week_start = 6) # EPI week starting on Saturday
x <- date2week(onset, week_start = 6, floor_day = TRUE) # truncate to just the weeks
table(x)
table(week2date(x))
```

## Installing the package

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:

```{r install2, eval = FALSE}
remotes::install_github("zkamvar/aweek")
```

Note that this requires the package *remotes* installed.


# Main Features

 - `date2week()` converts dates to a week format (YYYY-Www-d) that can start on 
   any day. 
 - `week2date()/as.Date()` does the backwards conversion from (YYYY-Www(-d)) to
   a numeric date. 
 - Dependencies only on R itself.

## Getting help online

Bug reports and feature requests should be posted on *github* using the [*issue*](http://github.com/reconhub/aweek/issues) system. All other questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via **pull requests**.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

