---
output: github_document
---

[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/rw)](https://cran.r-project.org/package=rw)

# Welcome to the *rw* package!

This package will convert dates to epiweeks, isoweeks, and all others in between
with minimal overhead.

```{r epiweek}

library(rw)
onset <- Sys.Date() + sample(-50:0, 100, replace = TRUE)
onset
date2week(onset, week_start = 1) # ISO weeks starting on Monday (default)
date2week(onset, week_start = 6) # EPI week starting on Saturday
x <- date2week(onset, week_start = 6, floor_day = TRUE) # rruncate to just the weeks
table(x)
table(week2date(x))
```


## Installing the package

To install the current stable, CRAN version of the package, type:

```{r install, eval = FALSE}
install.packages("rw")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:

```{r install2, eval = FALSE}
remotes::install_github("zkamvar/rw")
```

Note that this requires the package *remotes* installed.


# Main Features

 - `date2week()` converts dates to a week format (YYYY-Www-d) that can start on 
   any day. 
 - `week2date()/as.Date()` does the backwards conversion from (YYYY-Www(-d)) to
   a numeric date. 
 - Dependencies only on R itself.


# Resources

## Vignettes

An overview and examples of *rw* are provided in the vignettes:

...

## Websites

The following websites are available:

...

## Getting help online

Bug reports and feature requests should be posted on *github* using the [*issue*](http://github.com/reconhub/rw/issues) system. All other questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via **pull requests**.

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

