---
title: "aweek means 'any week'"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{aweek means 'any week'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Introduction
============

The day in which a week starts differs depending on context. For countries like
the UK, the first day of the week is the first working day, which is Monday.
This definition conforms with the ISO 8601 standard definition, but there are
examples of situations where the first day of the week is different:

 - The US CDC defines an "epiweek" which starts on a Sunday
 - In some regions, MSF will define a week that starts on Saturday

This package provides tools to convert dates to weeks and back where a week can
start on any day. You can use this package for any of the following:

 - convert date to week starting on any day
 - convert week back to date
 - convert week to week
 - create a factor of weeks that contains ordered levels that includes missing
   weeks. 

Converting dates to weeks with `date2week()`
==========================================

You can convert dates to weeks starting on any day by using `date2week()` with
the `week_start` argument. This argument can be a number from 1 to 7 representing
the ISO 8601 day of the week OR it can be a string representing the day of the
week in either an English locale or the locale defined on your computer.

```{r date2week}
library("aweek")

set.seed(2019-03-03)
dat <- as.Date("2019-03-03") + sample(-6:7, 10, replace = TRUE)
dat
# Use character days
print(w <- date2week(dat, week_start = "Sunday"))
# Use ISO 8601 days
print(w <- date2week(dat, week_start = 7))
```

Weekly aggregation 
------------------

There are times where you would want to aggregate your days into weeks, you can
do this by specifying `floor_day = TRUE` in `date2week()`. For example, here we
can show the individual weeks:

```{r date2week_floor}
print(wf <- date2week(dat, week_start = "Saturday", floor_day = TRUE))
table(wf)
```

Factors
-------

Weeks can be represented as factors, which is useful for tabulations across
weeks. You can use `factor = TRUE` in `date2week()` and it will automatically
fill in any missing days (`floor_day = FALSE`) or weeks (`floor_day = TRUE`)

```{r factors}
date2week(Sys.Date() + c(0, 15), week_start = 1, factor = TRUE)
date2week(Sys.Date() + c(0, 15), week_start = 1, factor = TRUE, floor_day = TRUE)
```


Weeks to weeks
--------------

You can also use `date2week()` to convert between different week definitions if
you have an `aweek` object:


```{r week2week}
w # week starting on Sunday
date2week(w, week_start = "wednesday")

# create a table with all days in the week
res <- as.data.frame(matrix("", nrow = 10, ncol = 7), stringsAsFactors = FALSE)
names(res) <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
for (i in names(res)) res[[i]] <- date2week(w, week_start = i)
as.data.frame(res)
```


Converting weeks to dates with `week2date()`
==========================================

If you receive data that contains week definitions, you can convert it back to 
a date if you know where the week starts.

```{r week2date}
week2date("2019-W10-1", week_start = "Sunday") # 2019-03-03
week2date("2019-W10-1", week_start = "Monday") # 2019-03-04
```

You can also use `as.Date()` and `as.POISXlt()` if you have an `aweek` object:

```{r asdate}
as.Date(w)
as.POSIXlt(w)
```


The `aweek` class 
===============

The `aweek` class is a character that contains the `week_start` attribute. This
attribute allows it to be easily converted back to a date without the user
needing to enter the start day every time. It can be created like so:

```{r aweek}
w <- "2019-W10-1"
attr(w, "week_start") <- 7 # monday
class(w) <- "aweek"
w
```

If you need to remove the class, you can just use `as.character()`:

```{r ascharacter}
as.character(w)
```

