---
title: "aweek means 'any week'"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{aweek means 'any week'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Introduction
============

The day in which a week starts differs depending on context. For countries like
the UK, the first day of the week is the first working day, which is Monday.
This definition conforms with the [ISO 8601 standard definition for the
begining of a week](http://en.wikipedia.org/wiki/ISO_week_date), but there are
examples of situations where the first day of the week is different:

 - [The US CDC defines an "MMWR" week which starts on a Sunday](https://wwwn.cdc.gov/nndss/document/MMWR_Week_overview.pdf)
 - In some regions, MSF will define a week that starts on Saturday

This package provides tools to convert dates to weeks and back where a week can
start on any day. You can use this package for any of the following:

 - convert date to week starting on any day
 - convert week numbers and years to dates
 - convert week to date
 - convert week to week
 - create a factor of weeks that contains ordered levels that includes missing
   weeks. 

Converting dates to weeks
=========================

You can convert dates to weeks starting on any day by using `date2week()` with
the `week_start` argument. This argument can be a number from 1 to 7 representing
the ISO 8601 day of the week OR it can be a string representing the day of the
week in either an English locale or the locale defined on your computer. The 
default of this argument is the value of `get_week_start()`, which is a thin
wrapper around `options("aweek.week_start", 1L)`. **Unless you have specified a
default `aweek.week_start` option with `set_week_start()`, this will always be 
set to 1 (Monday).**

> It is **highly recommended** that you set the default `aweek.week_start` either
> in the beginning of your Rscript, Rmarkdown document, or in your .Rprofile.

```{r date2week}
library("aweek")

set_week_start("Sunday") # setting the default week_start to Sunday

set.seed(2019-03-03)
dat <- as.Date("2019-03-03") + sample(-6:7, 10, replace = TRUE)
dat
print(w <- date2week(dat))
```

If you need a different day on the fly, you can supply an integer or character
day to the `week_start` argument.

```{r date2week_week_start}
# Use character days
date2week(dat, week_start = "Monday")
# Use ISO 8601 days
date2week(dat, week_start = 1)
```

If you want to save two extra keystrokes, you can also use the `as.aweek()`
method for dates, which wraps `date2week()`:

```{r as.aweek.Date}
as.aweek(dat, week_start = 1)
```

What you get back is an `aweek` class object. It can be converted back to a date
with either `as.Date()` or `week2date()`:

```{r date2week2date}
week2date(w)
as.Date(w)
```

The `aweek` class 
---------------

The result you see above is an object of class "aweek". The `aweek` class is a
character that contains the `week_start` attribute. This attribute allows it to
be easily converted back to a date without the user needing to enter the start
day every time. You can convert a character that matches the `YYYY-Www-d`
pattern to an `aweek` class object with `as.aweek()`:


```{r as.aweek.character}
x <- as.aweek("2019-W10-1")
x
```

> Under the hood, it checks the validity of the week string and then add the
> attribute and class:
> 
>     x <- "2019-W10-1"
>     attr(x, "week_start") <- 7 # Sunday 
>     class(x) <- "aweek"
>


If you need to remove the class, you can just use `as.character()`:

```{r ascharacter}
as.character(x)
```

Best practices
--------------

The `date2week()` function only checks that dates are in ISO 8601 (yyyy-mm-dd)
format before converting to weeks, *and otherwise assumes that the dates are
accurate* so it's strongly recommended to make sure your dates are in either
`Date` or `POISXt` format and accurate before converting to weeks. The 
[lubridate](https://cran.r-project.org/package=lubridate) can be used for this
purpose.

Use `set_week_start()` at the beginning of all your scripts to explicitly define
the day on wich your weeks start. This can be overridden if need be in specific
parts of your scripts. Otherwise, the default will be dependent on the value of
`getOption("aweek.week_start", 1L)`.

Because the `week_start` arguments default to `get_week_start()`, it's
recommended to specify `week_start` in `date2week()` and `week2date()` if you
don't have an `aweek` object.

Before you combine aweek objects, confirm that they are actually aweek objects
with `inherits(myObject, "aweek")`. 


Weekly aggregation 
------------------

There are times where you would want to aggregate your days into weeks, you can
do this by specifying `floor_day = TRUE` in `date2week()`. For example, here we
can show the individual weeks:

```{r date2week_floor}
print(wf <- date2week(dat, week_start = "Saturday", floor_day = TRUE))
table(wf)
```

If you convert this to date, then all the dates will represent the beginning
of the week:

```{r date2week_floor2date}
print(dwf <- week2date(wf))
weekdays(dwf)
```

If you want to aggregate your `aweek` objects after you created them, you can
always use the `trunc()` function:

```{r trunc}
w
trunc(w)
```


Factors
-------

Weeks can be represented as factors, which is useful for tabulations across
weeks. You can use `factor = TRUE` in `date2week()` and it will automatically
fill in any missing weeks.

```{r factors}
dat[1] + c(0, 15)
date2week(dat[1] + c(0, 15), week_start = 1, factor = TRUE)
```

> CAUTION: Do not use this with `floor_day = FALSE` as this will be removed
> in future versions.

Be careful when combining factors with other dates or aweek objects as they will
force the other objects to be truncated as well.

Weeks to weeks
--------------

You can use `change_week_start()` to convert between different week definitions
if you have an `aweek` object:


```{r week2week_wednesday}
w # week starting on Sunday
ww <- change_week_start(w, week_start = "wednesday") # same dates, starting on Wednesday
ww
identical(as.Date(w), as.Date(ww))
```

```{r week2week, R.options=list(width = 100)}
# create a table with all days in the week
d   <- as.Date("2019-03-03") + 0:6
res <- lapply(weekdays(d), function(i) date2week(d, week_start = i))
names(res) <- weekdays(d)
data.frame(res)
```

All of these columns contain the same dates:

```{r week2week2date, R.options=list(width = 100)}
data.frame(lapply(res, as.Date))
```

Combining `aweek` objects
-----------------------

You can add dates, aweek objects, or characters to aweek objects:

```{r caweekdate}
c(as.aweek("2010-W10-1"), 
  res$Sunday, 
  "2010-W12-1", 
  as.Date(res$Monday[1]) + 14)
```

However, you *can not* combine aweek objects with different `week_start` 
attributes.

```{r cweek2week_err}
try(c(res$Sunday[1], res$Wednesday[2], res$Friday[3]))
```

If you want to combine different aweek objects, you must first change their 
`week_start` attribute:

```{r cweekweek}
wed <- change_week_start(res$Wednesday, get_week_start())
fri <- change_week_start(res$Friday, get_week_start())
c(res$Sunday[1], wed[2], fri[3])
```

### Dates can be appended to aweek objects

Dates combined with aweek objects will will be automatically converted.

```{r add_dates}
c(res$Monday, as.Date("2019-04-03"))
```

### Add characters with caution

You can also add character representation of weeks, but be aware that **it is
assumed that these have the same `week_start` as the first object.**

```{r add_chars}
s <- c(res$Saturday, "2019-W14-3")
s
m <- c(res$Monday, "2019-W14-3")
m
```

**These will translate into different dates**

```{r char2date}
as.Date(s[7:8])
as.Date(m[7:8])
```

Working with weeks in data frames
=================================

You may encounter a situation where you have a merged data frame with weeks
starting on different days. This section will cover two situations where you
may have weeks as numbers and weeks as ISO-week strings. First we will create
our demonstration data that represents the same week with different `week_start`
attributes.

```{r date_a_frame}
# create a table with all days in the week
d    <- as.Date("2019-03-03") + 0:6
res  <- lapply(weekdays(d), function(i) date2week(d, week_start = i))
resn <- lapply(weekdays(d), function(i) date2week(d, week_start = i, numeric = TRUE))
datf <- data.frame(wday = rep(weekdays(d), each = 7), 
                   week = unlist(res), # note: unlist converts to character
                   week_number = unlist(resn),
                   year  = 2019,
                   stringsAsFactors = FALSE)
datf$day <- substring(datf$week, 10, 11)
head(datf, 10)
```

To get the weeks (numbers or strings) to aweek objects, you should use the
`start` argument to specify which day of the week they start on. Internally,
this translates the week to their corresponding dates and then to aweek objects
with the same `week_start` attribute (which defaults to `get_week_start()`).

weeks as numbers
------------

Most commonly, you will have weeks across data sets represented by numbers. 
These can be converted to aweek objects using the `get_aweek()` function 
and to dates using the `get_date()` function:

```{r get_aweek}
datf$aweek <- with(datf, get_aweek(week = week_number, year = year, day = day, start = wday))
datf$date  <- with(datf, get_date(week = week_number, year = year, day = day, start = wday))
head(datf, 10)
```

These functions are also useful for constructing weeks or dates on the fly if
you only have a week and a year:

```{r}
get_aweek(11, 2019)
get_date(11, 2019)
```

```{r remove_things, include = FALSE}
datf$aweek <- NULL
datf$date  <- NULL
```


weeks as characters
-------------------

If you have weeks formatted as ISO-week strings, then you can convert to aweek
objects using `as.aweek()`:

```{r date_a_frame_2}
datf$aweek <- with(datf, as.aweek(week, start = wday))
head(datf, 10)
str(datf)
```

We can tabulate them to see how they transformed:

```{r tabluate_data_frame, R.options = list(width = 100)}
print(with(datf, table(before = week, after = aweek)), zero.print = ".")
```



Converting weeks to dates
=========================

If you receive data that contains week definitions, you can convert it back to 
a date if you know where the week starts.

```{r week2date}
week2date("2019-W10-1", week_start = "Sunday") # 2019-03-03
week2date("2019-W10-1", week_start = "Monday") # 2019-03-04
```

If you have an `aweek` object, however, it will use the `week_start` attribute
defined in the object, even if the default `week_start` attribute is different:

```{r week2date_aweek}
set_week_start("Monday") # Set the default week_start to ISO week
get_week_start(w)        # show the default week_start for w
week2date(w)
identical(week2date(w), dat)               # TRUE
identical(week2date(as.character(w)), dat) # FALSE
```


You can also use `as.Date()` and `as.POISXlt()` if you have an `aweek` object:

```{r asdate}
as.Date(w)
as.POSIXlt(w)
```

