---
title: "aweek means 'any week'"
date: "`r Sys.Date()`"
output:
   rmarkdown::html_vignette:
     toc: true
     toc_depth: 2
vignette: >
  %\VignetteIndexEntry{aweek means 'any week'}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Introduction
============

The day in which a week starts differs depending on context. For countries like
the UK, the first day of the week is the first working day, which is Monday.
This definition conforms with the [ISO 8601 standard definition for the
begining of a week](http://en.wikipedia.org/wiki/ISO_week_date), but there are
examples of situations where the first day of the week is different:

 - [The US CDC defines an "MMWR" week which starts on a Sunday](https://wwwn.cdc.gov/nndss/document/MMWR_Week_overview.pdf)
 - In some regions, MSF will define a week that starts on Saturday

This package provides tools to convert dates to weeks and back where a week can
start on any day. You can use this package for any of the following:

 - convert date to week starting on any day
 - convert week back to date
 - convert week to week
 - create a factor of weeks that contains ordered levels that includes missing
   weeks. 

Converting dates to weeks with `date2week()`
==========================================

You can convert dates to weeks starting on any day by using `date2week()` with
the `week_start` argument. This argument can be a number from 1 to 7 representing
the ISO 8601 day of the week OR it can be a string representing the day of the
week in either an English locale or the locale defined on your computer.

```{r date2week}
library("aweek")

set.seed(2019-03-03)
dat <- as.Date("2019-03-03") + sample(-6:7, 10, replace = TRUE)
dat
# Use character days
print(w <- date2week(dat, week_start = "Sunday"))
# Use ISO 8601 days
print(w <- date2week(dat, week_start = 7))
```

What you get back is an `aweek` class object. It can be converted back to a date
with either `as.Date()` or `week2date()`:

```{r date2week2date}
week2date(w)
```


The `aweek` class 
---------------

The result you see above is an object of class "aweek". The `aweek` class is a
character that contains the `week_start` attribute. This attribute allows it to
be easily converted back to a date without the user needing to enter the start
day every time. It can be created like so: 

```{r aweek}
x <- "2019-W10-1"
attr(x, "week_start") <- 7 # monday
class(x) <- "aweek"
x
class(x)
```

If you need to remove the class, you can just use `as.character()`:

```{r ascharacter}
as.character(x)
```


Weekly aggregation 
------------------

There are times where you would want to aggregate your days into weeks, you can
do this by specifying `floor_day = TRUE` in `date2week()`. For example, here we
can show the individual weeks:

```{r date2week_floor}
print(wf <- date2week(dat, week_start = "Saturday", floor_day = TRUE))
table(wf)
```

If you convert this to date, then all the dates will represent the beginning
of the week:

```{r date2week_floor2date}
print(dwf <- week2date(wf))
weekdays(dwf)
```

Factors
-------

Weeks can be represented as factors, which is useful for tabulations across
weeks. You can use `factor = TRUE` in `date2week()` and it will automatically
fill in any missing days (`floor_day = FALSE`) or weeks (`floor_day = TRUE`)

```{r factors}
date2week(Sys.Date() + c(0, 15), week_start = 1, factor = TRUE)
date2week(Sys.Date() + c(0, 15), week_start = 1, factor = TRUE, floor_day = TRUE)
```


Weeks to weeks
--------------

You can also use `date2week()` to convert between different week definitions if
you have an `aweek` object:


```{r week2week}
w # week starting on Sunday
date2week(w, week_start = "wednesday")

# create a table with all days in the week
res <- as.data.frame(matrix("", nrow = 10, ncol = 7), stringsAsFactors = FALSE)
names(res) <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
for (i in names(res)) res[[i]] <- date2week(w, week_start = i)
as.data.frame(res)
```


Converting weeks to dates with `week2date()`
==========================================

If you receive data that contains week definitions, you can convert it back to 
a date if you know where the week starts.

```{r week2date}
week2date("2019-W10-1", week_start = "Sunday") # 2019-03-03
week2date("2019-W10-1", week_start = "Monday") # 2019-03-04
```

If you have an `aweek` object, however, it will use the `week_start` attribute
defined in the object:

```{r week2date_aweek}
week2date(w)
```


You can also use `as.Date()` and `as.POISXlt()` if you have an `aweek` object:

```{r asdate}
as.Date(w)
as.POSIXlt(w)
```

